Visualizing probability distributions across bivariate nonlinear temporal granularities

### 2020-02-24 main.Rmd
content cut iteration 2
 - def bottom gran
 + def calendar

\begin{definition}[Finer than]\label{def:finerthan}
A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$.
\end{definition}.

@@

These definitions are illustrated in \autoref{fig:groups-finer}, where both day $\trianglelefteq$ week, and day $\preceq$ week, hold true. However, weekends are finer than weeks but do not group into weeks. And weekends neither groups into or is finer than public holidays. Consider another example where day $\trianglelefteq$ month. This relationship is not fully described until we associate it with periodicity. Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored. The period becomes 400 years with the inclusion of leap years and all their exceptions.

```{r groups-finer, echo=FALSE, out.width = "100%", fig.cap="Illustrations of groups into and finer than relationships. Weekends are finer than weeks but do not group into weeks. And weekends neither groups into or finer than public holidays"}
knitr::include_graphics("Figs/groupsandfiner.png")
```

@@

A granularity $H$ which is periodical with respect to $G$ is specified by:
(i) the $R$ sets of indexes of $G$, ${S_0,\dots,S_{R-1}}$ describing the granules of $H$ within one period;
(ii) the value of $P$;
(iii) the indexes of first and last granules in $H$, if their value is not infinite.
Then, if ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
$H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$. Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy].

@@

\begin{definition}[Groups quasi-periodically]\label{def:quasi}
A granularity $H$ is quasi-periodical with respect to a granularity $G$ if:
(1) $G \trianglelefteq H$; and
(2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$.
\end{definition}

Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.

@@

Consider the Gregorian calendar without leap years. Then each month is a grouping of the same number of days each year. The period is $P = 365$ and the number of granules of "months" in each period is given by $R=12$. Thus, we could say days groups periodically into months. Now consider the Gregorian calendar with leap years, so that days group quasi-periodically into months. The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s would be the time domain corresponding to $29^{\text{th}}$ February of any year.

\begin{definition}[Bottom granularity]\label{def:B}
Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set.
\end{definition}

Here, g-rel could be any relationship defined between granularities like "groups into", "finer than" or "groups periodically". Given the set of all granularities defined over the time domain $(\mathbb{Z}; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$.

<!-- Most practical problems seem to require only a granularity system containing a set of time granularities which are all periodical (or quasi-periodical) with respect to a basic granularity. The assumption is there should be one bottom granularity and all the other granularities are generated from the bottom granularity by calendar algebraic operations. -->

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate new granularities recursively from the bottom up. The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations. The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity. For example, consider a calendar with only two linear granularities: "minute" and "hour", where "hour" is generated by grouping every $60$ "minutes". The granule-oriented operations do not change the granules of a granularity, but rather select which granules should remain in the new granularity. For example, one can define the granularity "Monday" and hence select only Mondays from linear granularity "day".

<!-- It might be the case that we do not always have integers as the label sets of granularities.  -->

@@

The notion of a labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra. A labeled granularity is a pair $(L, G)$, where $L$ is a subset of the integers, and $G$ is a mapping from $L$ to the subsets of the time domain such that:

  1. if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$; and
  2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.

$L$ is called the label set of $G$. When $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}. However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules. Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.

Some relevant grouping-oriented operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.

The grouping operation:
: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer. The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity. More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$,
$$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
For example:
\begin{align*}
  \text{minute} &= \text{Group}_{60}(\text{second});\\
  \text{hour}   &= \text{Group}_{60}(\text{minute});\\
  \text{day}    &= \text{Group}_{24}(\text{hour});\\
  \text{week}   &= \text{Group}_{7}(\text{day});
\end{align*}
where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise. Otherwise, repeated operations of grouping might result in a granule of "resultant" granularity "day" or "week", which is different from a time-stamped "day" or "week".

@@

For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.  Let us consider a granularity "pseudomonth" which computes months only for non-leap years.
\begin{align*}
\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \\
  &~~\text{Alter}^{12}_{9,-1}(\text{day}, \\
  &~~~~\text{Alter}^{12}_{6,-1}(\text{day}, \\
  &~~~~~~\text{Alter}^{12}_{4,-1)}(\text{day}, \\
  &~~~~~~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \\
  &~~~~~~~~~~\text{Group}_{31}(\text{day})))))),
\end{align*}
where the granularity pseudomonth is generated by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].

@@

In general, any circular granularity relating two linear granularity, none of which are bottom granularity\todo{Why not?} can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$. Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping. It is possible that none of the two linear granularities are bottom granularities. But the representation of the resultant circular granularity will be a function of the index set. The bottom granularity is assumed to be minutes.

@@

Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
These are called "structurally" empty combinations. Let us take a specific example, where $C_1$ maps index set to day-of-month and $C_2$ maps index set to week-of-month. Here $C_1$ can have 31 levels/categories while $C_2$ can have 5 categories. There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of day-of-month and week-of-month. Many of these are empty. For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month. These are structurally empty sets in that it is impossible for them to have any observations.

@@

Secondly, empty combinations can turn up due to differences in event location or duration in a calendar. These are called "event-driven" empty combinations. Again, let us consider a specific example to illustrate this. Let $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay. Here $C_1$ can have 7 levels while $C_2$ can have 2 levels. So there are 14 sets $S_{ij}$ corresponding to the possible combinations of day-of-week and WorkingDay/NonWorkingDay. While potentially all of these can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these combinations will probably have very few observations. For example, there are few if any public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.

@@

Thirdly, empty combinations can be a result of how granularities are constructed. Let $C_1$ maps index set to Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month. Then the days denoting weekends in a month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed. This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does. Hence, they are referred to as "build-based" empty combinations.

@@

Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation. With huge amount of data being available, mean, median or any one summary statistic might not be enough to understand the different properties of the measured variable. Instead the distribution could be summarized using some summary statistics together, each of which can highlight different properties of the data. Tukey's five number summary, letter values or quantiles can serve as examples of distributional summaries. Another way of summarizing could be estimating the probability density function (continuous) or probability mass function (dicrete). The approach can be parametric or non-parametric. A non-parametric approach makes less rigid assumptions about the distribution of the observed data making the data speak more for itself. These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses. Several forms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:

@@

where $h$ is the smoothing parameter or the bandwidth. The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by:

<!-- $Q$, $(q \equiv  Q')$, which is the quantile density function -->
<!-- $ Q_{n} (u) = \sum_{i=1}^ n X_{(i)} I((i-1)/n <u \leq i/n)$\\ -->
<!--             $= X_(1) + \sum_{i=1}^ n (X_{(i)} - X_{(i-1)}) I((i-1)/n < u)$ -->

<!-- and most widely used definition of sample quantile is given in  by: -->

<!-- $$\hat{Q_{i}(p)} = (l - \gamma)X_{(j)} + \gamma X_{(j + 1)}$$ where,  $(j - m)/n \leq p < (j - m + 1)/n$,  -->
<!-- for some $m \in R$ and $0 < \gamma < 1$.  $\hat{Qi(p)}$ is the ith sample quantile  -->

\begin{equation}
\hat{q(u)} = \sum_{i=2}^{n}(X_{(i)} - X_{(i-1)})K(u - (i-1)/n)) - X_{(n)}K(u-1) + X_{(1)}K(u)
\end{equation}

where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$.

 <!-- Both pdf and pmf are concerned with point probabilities of the random variable. Suppose we have observations which are realizations of univariate random variables, $X_1, X_2, \dots, X_n \sim F$, where $F$ denotes an unknown cumulative distribution function and $X_1, X_2, \dots, X_n$ are i.i.d. -->

<!-- Instead of assuming a parametric model for the distribution (e.g. a Normal distribution with unknown expectation and variance), a non-parametric estimation assumes that the density exists and is suitably smooth (e.g. differentiable). It is then possible to estimate the u2nknown density function f(·). -->

<!-- The histogram is the oldest and most popular density estimator method. The form of the histogram depends very much on the tuning parameters class width and origin. The naive estimator was introduced by Fix and Hodges (1951) and first discussed by Rosenblatt(1956). Whittle (1958) formulated the general weight function class of estimators, while the orthogonal series estimator was introduced by Cencov (1962). The nearest neighbor estimate considered by Loftsgaarden and Quesenbury (1965), while the variable Kernel method is due to Brieman, Meisel and Purcell (1977). There are quite a few other approaches to density estimation than the kernel estimators, but the fixed bandwidth kernel estimators are used predominantly because of their simplicity. -->

<!-- The inverse $Q = F^{-1}$, which is the quantile function can be defined as $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$.  -->

The MSE of the estimators are given as follows:

\begin{equation}\label{eq:MSE}
\begin{split}
MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\
MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa
\end{split}
\end{equation}

@@

@@

Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization. The basic plot choice for our data structure is the one that can display densities. In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated using Kernel density estimates or some descriptive statistics. Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set. The descriptive statistics based methods do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former. However, they avoid clutter and help us to focus on some specific properties of the data. Also, the summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.

<!--  However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data. -->

We will discuss few conventional and recent ways to plot distributions using both of these methods.

<!-- Non-parametric density based plots can show the entire distribution, unlike discretizing and showing only parts of it. These displays depend on the specific estimation procedure (e.g., kernel density estimate)  -->

@@

Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers. These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles. Due to wide spread popularity and simplicity in implementation, a number of variations are proposed to the original one which provides alternate definitions of quantiles, whiskers, fences and outliers. Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance. However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy. Also, the number of outliers is large for larger data set since number of outliers is proportional to the number of observations.

<!-- Description and lit review of letter-value plots -->
The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail. These are particularly useful for large data sets. “outliers” in letter value plots are those observations beyond the most extreme letter value. The letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.

@@

Traditional ways to visualize kernel densities include violin plots [@Hintze1998-zi]. The shape of the violin represents the density estimate of the variable. The more data points in a specific range, the larger the violin is for that range. Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution.

@@

VIC %>%
  create_gran("month_year") %>%
  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>%
  prob_plot("month_year","day_week",
            response = "Demand",
            plot_type = "lv") +
  ylab("Electricity Demand [KWh]") +
  ggtitle("(a) Letter value plot")

VIC %>%
  create_gran("day_week") %>%
  filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>%
  prob_plot("day_week","hour_day",
            response = "Demand",
            plot_type = "quantile",
            symmetric = FALSE) +
  ylab("Electricity Demand [KWh]") +
  ggtitle("(b) Decile plot")

VIC %>%
  create_gran("day_month") %>%
  filter(day_month %in% c(1, 15, 29, 31)) %>%
  prob_plot("day_month",
                "day_year",
                response = "Demand",
                plot_type = "boxplot")+
  ylab("Electricity Demand [KWh]") +
  scale_x_discrete(breaks = seq(0, 366, 60)) +
  theme(legend.position = "bottom",
        strip.text = element_text(size = 7, margin = margin())) +
  ggtitle("(c) Box plot")

VIC %>%
    create_gran("week_month") %>%
  filter(week_month %in% c(1, 2, 4)) %>%
   prob_plot("week_month",
                "day_month",
                response = "Demand",
                plot_type = "violin", alpha = 0.3) + ylab("") +
  xlab("Days of the Month") +
  theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) +
  scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
  ggtitle("(d) Violin plot")

>>>>

<!-- CONTENT CUT -->
<!-- \begin{definition}[Finer than]\label{def:finerthan} -->
<!-- A linear granularity $G$ is finer than a linear granularity $H$, denoted $G \preceq H$, if for each index $i$, there exists an index $j$ such that $G(i) \subset H(j)$. -->
<!-- \end{definition}. -->

@@

<!-- CONTENT CUT -->
<!-- These definitions are illustrated in \autoref{fig:groups-finer}, where both day $\trianglelefteq$ week, and day $\preceq$ week, hold true. However, weekends are finer than weeks but do not group into weeks. And weekends neither groups into or is finer than public holidays. Consider another example where day $\trianglelefteq$ month. This relationship is not fully described until we associate it with periodicity. Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored. The period becomes 400 years with the inclusion of leap years and all their exceptions. -->

For example, day $\trianglelefteq$ week since every granule of "week" is the union of some set of granules of "day". Consider another example where day $\trianglelefteq$ month. This relationship however is incomplete without its association to periodicity. Each month is a grouping of the same number of days over years, hence the period of the grouping (day, month) is one year, if leap years are ignored. The period becomes 400 years with the inclusion of leap years and all their exceptions.

<!-- CONTENT CUT -->
<!-- # ```{r groups-finer, echo=FALSE, out.width = "100%", fig.cap="Illustrations of groups into and finer than relationships. Weekends are finer than weeks but do not group into weeks. And weekends neither groups into or finer than public holidays"} -->
<!-- knitr::include_graphics("Figs/groupsandfiner.png") -->
<!-- ``` -->

@@

<!-- CONTENT CUT -->
<!-- A granularity $H$ which is periodical with respect to $G$ is specified by: -->
<!-- (i) the $R$ sets of indexes of $G$, ${S_0,\dots,S_{R-1}}$ describing the granules of $H$ within one period; -->
<!-- (ii) the value of $P$; -->
<!-- (iii) the indexes of first and last granules in $H$, if their value is not infinite. -->

If ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
$H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$. Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy]. These are called quasi-periodic relationships. Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.

@@

<!-- CONTENT CUT -->
<!-- \begin{definition}[Groups quasi-periodically]\label{def:quasi} -->
<!-- A granularity $H$ is quasi-periodical with respect to a granularity $G$ if: -->
<!-- (1) $G \trianglelefteq H$; and -->
<!-- (2) there exist a set of intervals $E_1,\dots,E_z$ (the granularity exceptions) and positive integers $R$, $P$, where $R$ is less than the minimum of the number of granules of $H$ between any two exceptions, such that for all $i \in \mathbb{Z}$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where $E$ is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P)$. -->
<!-- \end{definition} -->

@@

For example, in a Gregorian calendar without leap years each month is a grouping of the same number of days each year. The period is $P = 365$ and the number of granules of "months" in each period is given by $R=12$. Thus, we could say days group periodically into months. In a Gregorian calendar with leap years, days group quasi-periodically into months. The value of $P$ and $R$ would still be 365 and 12 respectively, while the <!--$E_i$s--> exceptions would be the time domain corresponding to $29^{\text{th}}$ February of any year.

<!-- CONTENT CUT -->
<!-- \begin{definition}[Bottom granularity]\label{def:B} -->
<!-- Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity B in the set is a bottom granularity with respect to g-rel, if B g-rel H for each granularity H in the set. -->
<!-- \end{definition} -->

<!-- Here, g-rel could be any relationship defined between granularities as defined in @Bettini1998-ed. Given the set of all granularities defined over the time domain $(\mathbb{Z}; <)$, and the granularity relationship $\trianglelefteq$ (groups into), the granularity mapping each index into the corresponding instant (same integer number as the index) is a bottom granularity with respect to $\trianglelefteq$. -->

A discrete time model often uses a fixed smallest granularity also known as _*bottom granularity_* (@Bettini1998-ed) so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _*chronons*_ (@Jensen1998-qn). Chronons may be grouped into granules and a linear granularity is a non-overlapping mapping of 
granules to subsets of the time domain (see @Dyreson2000-mr). A system of multiple granularities in lattice structures is referred to as a calendar.

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a “bottom” granularity and calendar algebra operations are designed to generate new granularities recursively from the bottom up.

<!-- CONTENT CUT -->
<!-- The calendar algebra consists of two kinds of operations: grouping-oriented and granule-oriented operations. The grouping-oriented operations combine certain granules of a granularity together to form the granules of the new granularity. For example, consider a calendar with only two linear granularities: "minute" and "hour", where "hour" is generated by grouping every $60$ "minutes". -->
<!-- The granule-oriented operations do not change the granules of a granularity, but rather select which granules should remain in the new granularity. For example, one can define the granularity "Monday" and hence select only Mondays from linear granularity "day". -->

@@

In addition to the integer index, each granule could be associated with a textual representation which is called the label. The label can be mapped to index for input and the index can later be mapped to this label for output. When the label set is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}. Some calendar operations will only work on full-integer labeled granularities, while others can be implemented using more flexible labelings.

<!-- CONTENT CUT -->
<!-- The notion of a labeled granularity is introduced in @Bettini2000-qk to simplify the calendar algebra. A labeled granularity is a pair $(L, G)$, where the label set $L$ is a subset of the integers instead of the set of integers and -->
<!-- follows the conditions (1) and (2) in \ref{def:linear} where $i$, $j$, and $k$ are in $L$. When the label set $L$ is exactly the integers, the granularities are called "full-integer" labeled granularity, such as in Definition \ref{def:linear}.<!-- However, the set $L$ of labels can be an arbitrary subset of (possibly non contiguous) integers and could be used to identify granules.-->  -->

<!-- $G$ is a mapping from $L$ to the subsets of the time domain such that: -->

<!--   1. if $i<j$ , where $i$ and $j$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then each element in $G(i)$ is less than all the elements of $G(j)$; and -->
<!--   2. if $i<k<j$ , where $i$, $j$, and $k$ are in $L$, and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty. -->

Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} to define circular and aperiodic granularities.

The grouping operation:
: Let $G_1$ be a full-integer labeled granularity, and $m$ a positive integer. The grouping operation Group$_m(G)$ generates a new granularity $G_2$, by partitioning the granules of $G_1$ into $m$-granule groups and making each group a granule of the resulting granularity. More precisely, $G_2 = \text{Group}_m(G_1)$ is the full-integer labeled granularity such that for each integer $i$,
$$G_2(i) = \bigcup\limits_{j = (i-1)m+1}^{im} G_1(j).$$
Example: $\text{minute} = \text{Group}_{60}(\text{second})$; $\text{hour} = \text{Group}_{60}(\text{minute})$

<!-- CONTENT CUT -->
<!-- where $\text{second}(1)$ should coincide with $\text{minute}(1)$ and likewise. Otherwise, repeated operations of grouping might result in a granule of "resultant" granularity "day" or "week", which is different from a time-stamped "day" or "week". -->

@@

For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.  Let us consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].

\begin{align*}
\text{pseudomonth} = &\text{Alter}^{12}_{11,-1}(\text{day}, \text{Alter}^{12}_{9,-1}(\text{day}, \\
  &~~\text{Alter}^{12}_{6,-1}(\text{day}, \text{Alter}^{12}_{4,-1)}(\text{day}, \\
  &~~~\text{Alter}^{12}_{(2,-3)}(\text{day}, \text{Group}_{31}(\text{day})))))),
\end{align*}

@@

In general, any circular granularity relating two linear granularity can be expressed as $C_{(G, H)}(z) = L(\lfloor z/P(B,G) \rfloor\mod P(G,H))$, where linear granularity $H$ is periodic with respect to linear granularity $G$ with regular mapping such that the number of granules of $G$ in each granule of $H$ is $P(G,H)$. Table \ref{tab:definitions} shows representation of circular granularities $C_i$ relating two linear granularities with $P_i$ being the number of granules of the finer granularity in each granule of the coarser granularity and $L_i$ is the associated label mapping. It is possible that none of the two linear granularities are bottom granularities. But the representation of the resultant circular granularity will be a function of the index set. The bottom granularity is assumed to be minutes.

@@

Firstly, empty combinations can arise due to the structure of the calendar or hierarchy.
These are called "structurally" empty combinations. Let us take a specific example, where $C_1$ be day-of-month and $C_2$ be week-of-month. Here $C_1$ can have 31 levels/categories while $C_2$ can have 5 categories. There will be $31\times 5=155$ sets $S_{ij}$ corresponding to the possible combinations of day-of-month and week-of-month. Many of these are empty. For example $S_{1,5}$, $S_{21,2}$, etc.This is also intuitive since the first day of the month can never correspond to fifth week of the month. These are structurally empty sets in that it is impossible for them to have any observations.

@@

Secondly, empty combinations can turn up due to differences in event location or duration in a calendar. These are called "event-driven" empty combinations. Again, let us consider a specific example to illustrate this. Let $C_1$ be day-of-week and $C_2$ be WorkingDay/NonWorkingDay. Here $C_1$ can have 7 levels while $C_2$ can have 2 levels. So there are 14 sets $S_{ij}$ corresponding to the possible combinations of $C_1$ and $C_2$. While potentially all of these can be non-empty (it is possible to have a public holiday on any day-of-week), in practice many of these combinations will probably have very few observations. For example, there are few (if any) public holidays on Wednesdays or Thursdays in any given year in Melbourne, Australia.

@@

Thirdly, empty combinations can be a result of how granularities are constructed. Let $C_1$ be Business-days, which are days from Monday to Friday except holidays and $C_2$ is day-of-month. Then the days denoting weekends in a month would not correspond to any Business days and would have missing observations due to the way the granularities are constructed. This is different from structurally empty combinations because structure of the calendar does not lead to these missing combinations, but the construction of the granularity does. Hence, they are referred to as "build-based" empty combinations.

@@

Summarizing properties of the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation. With huge amount of data being available, mean, median or any one summary statistic might not be enough to understand the different properties of the measured variable. Instead the distribution could be summarized using some summary statistics together, each of which can highlight different properties of the data. Tukey's five number summary, letter values or quantiles can serve as examples of distributional summaries. Another way of summarizing could be estimating the probability density function (continuous) or probability mass function (dicrete). The approach can be parametric or non-parametric. A non-parametric approach makes less rigid assumptions about the distribution of the observed data making the data speak more for itself. These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses. Several forms of  non-parametric density estimators <!--and quantile estimators--> exist in the literature, the most commonly used form in @Silverman1986-cl<!--,  @Jones1992-pj)--> is as follows:

@@

where $h$ is the smoothing parameter or the bandwidth. The MSE of the estimator is given by:

\begin{equation}\label{eq:MSE}
\begin{split}
MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\
\end{split}
\end{equation}

@@

<!-- CONTENT CUT -->
<!-- The sample quantile estimator  $\hat{q(u)}$ with Kernel $K$ is given by: -->

<!-- \begin{equation} -->
<!-- \hat{q(u)} = \sum_{i=2}^{n}(X_{(i)} - X_{(i-1)})K(u - (i-1)/n)) - X_{(n)}K(u-1) + X_{(1)}K(u) -->
<!-- \end{equation} -->

<!-- where $Q(u)= F^{-1}(u) = inf({x: F(x) >u)}$, $0<u< 1$ and  $q \equiv  Q'$. -->
<!-- MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa -->

@@

Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization. The basic plot choice for our data structure is the one that can display densities. In Section \ref{sec:measured-variable}, it is discussed that these densities could be estimated using Kernel density estimates or some descriptive statistics.
We will discuss few conventional and recent ways to plot distributions using both of these methods.

<!-- CONTENT CUT -->
<!-- Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set. The descriptive statistics based methods do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former. However, they avoid clutter and help us to focus on some specific properties of the data. Also, the summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration. -->

@@

Most commonly used displays include the box plots [@Tukey1977-jx] which convey statistical features such as the median, quartile boundaries, hinges, whiskers and extreme outliers.<!--These displays are based on descriptive statistics, for example, $([n/2] + 1)/2^{\text{th}}$ and $(n + 1 - ([n/2] + 1)/2)^{\text{th}}$ order statistics are taken as estimates of the quartiles.--> Due to wide spread popularity and simplicity in implementation, a number of variations are proposed which provides alternate definitions of quantiles, whiskers, fences and outliers. Notched box plots [@Mcgill1978-hg] has box widths proportional to the number of points in the group and display confidence interval around medians aims to overcome some drawbacks of box plots.<!-- Challenge of boxplot --> The standard box plot and all of it's variations are helpful to get an idea of the distribution at a glance. However, for data less than 1000 observations, detailed estimates of tail behavior beyond the quartiles are not trustworthy. Also, the number of outliers is large for larger data sets.

<!-- Description and lit review of letter-value plots -->
The letter-value box plot [@Hofmann2017-sg] was designed to adjust for number of outliers proportional to the data size and display more reliable estimates of tail and are particularly useful for large data sets. “outliers” in letter value plots are those observations beyond the most extreme letter value. The letter values are shown till the depths where the letter values are reliable estimates of their corresponding quantiles and hence might lead to a lot of letter values being shown, leading to overload of information in one plot.

@@

Traditional ways to visualize kernel densities include violin plots [@Hintze1998-zi]. The shape of the violin represents the density estimate of the variable. The more data points in a specific range, the larger the violin is for that range. Adding two density plots gives a symmetric plot which makes it easier to see the magnitude of the density and compare across categories, enabling easier detection of clusters or bumps within a distribution. It becomes less useful when there are too many levels or categories to display.

@@

VIC %>%
  create_gran("month_year") %>%
  filter(month_year %in% c("Jan", "Mar", "July", "Dec")) %>%
  prob_plot("month_year", "day_week",
    response = "Demand",
    plot_type = "lv"
  ) +
  ylab("Electricity Demand [KWh]") +
  ggtitle("(a) Letter value plot")

VIC %>%
  create_gran("day_week") %>%
  filter(day_week %in% c("Mon", "Tue", "Sat", "Sun")) %>%
  prob_plot("day_week", "hour_day",
    response = "Demand",
    plot_type = "quantile",
    symmetric = FALSE
  ) +
  ylab("Electricity Demand [KWh]") +
  ggtitle("(b) Decile plot")

VIC %>%
  create_gran("day_month") %>%
  filter(day_month %in% c(1, 15, 29, 31)) %>%
  prob_plot("day_month",
    "day_year",
    response = "Demand",
    plot_type = "boxplot"
  ) +
  ylab("Electricity Demand [KWh]") +
  scale_x_discrete(breaks = seq(0, 366, 60)) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 7, margin = margin())
  ) +
  ggtitle("(c) Box plot")

VIC %>%
  create_gran("week_month") %>%
  filter(week_month %in% c(1, 2, 4)) %>%
  prob_plot("week_month",
    "day_month",
    response = "Demand",
    plot_type = "violin", alpha = 0.3
  ) + ylab("") +
  xlab("Days of the Month") +
  theme(legend.position = "bottom", strip.text = element_text(size = 7, margin = margin())) + scale_x_discrete(breaks = seq(0, 31, 5)) +
  scale_y_continuous(breaks = seq(2000, 9000, 2000)) +
  ggtitle("(d) Violin plot")
